Approximation as Differentiation

### 2025-05-31 main/higher-order.tex
Note on reindexing functor for f, which is just the image of f in the indexed category [-,C].
Start on typing rules.
WIP.
Some citations on categories for types.
Defining mu-polynomial functors.

\vspace{-3mm}
\begin{align*}
\lambda_{X,Y,Z}(f)(x) &= g \quad\text{with $g = f \comp (x,-)$ and $\partial g(x) = \reindex{\partial f}{(x,-)}$} \\
\partial\lambda_{X,Y,Z}(f)(x) &= \prodFam{\{\partial f(x,y) \comp \biinj_{\partial X(x)}\}_{y \in Y}}
\end{align*}
\noindent Here $(x,-): Y \to X \times Y$ is the ``reindexing'' function that pairs any $y$ with a fixed $x$,
and for any $y$ $\biinj_{\partial X(x)}$ denotes the canonical injection $\partial X(x) \to \partial X(x)
\biprod \partial Y(y)$.

\todo{
\begin{enumerate}
  \item $\FinVect$ and $\LatGal$ both have biproducts but neither has all small products
  \item $\PSh(\cat{C})$ has all small products (inherited from $\Set$), but doesn't automatically inherit
  biproducts from $\cat{C}$
  \item however, $\PSh_{\CMon}(\cat{C})$ does indeed have biproducts if $\cat{C}$ does
  \item thus $\Fam(\PSh_{\CMon}(\FinVect))$ and $\Fam(\PSh_{\CMon}(\LatGal))$ are both Cartesian closed
  \item also need that $\CMon$-enriched Yoneda embedding preserves biproducts \todo{check: because this is how we will interpret base types? E.g.~$\sem{\mathsf{Bool}} = \Hom{\cat{C}}{-}{2}$ assuming $\cat{C}$ provides a suitable object 2}
  \end{enumerate}
}

@@

>>>>

\vspace{-4mm}
\begin{align*}
\lambda_{X,Y,Z}(f)(x) &= g \quad\textit{with $g = f \comp (x,-)$ and $\partial g(x) = \reindex{\partial f}{(x,-)}$} \\
\partial\lambda_{X,Y,Z}(f)(x) &= \prodFam{\{\partial f(x,y) \comp \biinj_{\partial X(x)}\}_{y \in Y}}
\end{align*}
\noindent Here $(x,-): Y \to X \times Y$ is the ``reindexing'' function that pairs any $y$ with a fixed $x$,
and for any $y$ $\biinj_{\partial X(x)}$ denotes the canonical injection $\partial X(x) \to \partial X(x)
\biprod \partial Y(y)$.

\todo{
\begin{enumerate}
  \item $\FinVect$ and $\LatGal$ both have biproducts but neither has all small products
  \item $\PSh(\cat{C})$ has all small products (inherited from $\Set$), but doesn't automatically inherit
  biproducts from $\cat{C}$
  \item however, $\PSh_{\CMon}(\cat{C})$ does indeed have biproducts if $\cat{C}$ does
  \item thus $\Fam(\PSh_{\CMon}(\FinVect))$ and $\Fam(\PSh_{\CMon}(\LatGal))$ are both Cartesian closed
  \item also need that $\CMon$-enriched Yoneda embedding preserves biproducts \todo{check: because this is how
  we will interpret base types? E.g.~$\sem{\mathsf{Bool}} = \Hom{\cat{C}}{-}{2}$ assuming $\cat{C}$ provides a
  suitable object 2?}
  \end{enumerate}
}

@@

\subsection{Higher-order language}
\todo{Introduce earlier?}

We introduce a standard total functional languages with a (reasonably) expressive type
system~\cite{crole94,pitts01,santocanale02}.

\subsubsection{Syntax}

\input{fig/mu-polynomial}

\begin{definition}[$\mu$-polynomial]
Suppose $\cat{C}$ a category with finite coproducts and finite products. Define $\muPoly$
to be the smallest subcategory of $\Cat$ closed under the rules in \figref{mu-polynomial}.
\end{definition}

\input{fig/syntax}

\figrefTwo{syntax}{typing} give the syntax and typing rules of the higher-order language.

\subsubsection{Semantics}

\input{fig/typing}
